---
title: "R Notebook"
output: html_notebook
---

```{r}
library(readxl)
library(readr)
library(dplyr)
library(here)
library(tidyverse)
library(stringr)
```

# Read in xlsx file

```{r}
sheet1 <- read_xlsx(here("raw_data/boing-boing-candy-2015.xlsx"))
sheet1
```

# Inspect the data i.e. dimensions, variable types, variable names, etc.?

```{r}
dim(sheet1)
colnames(sheet1)
glimpse(sheet1)
```

# Tidy data - long format

```{r}
sheet1_long <- sheet1 %>%
  pivot_longer(cols = starts_with("["),  #converting rows to columns 
               names_to = "candy",
               values_to = "rating")

#check dimensions, column names
dim(sheet1_long)
colnames(sheet1_long)
```

#Omitting columns that are not required

```{r}
sheet1_long <- sheet1_long %>%
  select(-c(4:29))

colnames(sheet1_long)
dim(sheet1_long)
```


```{r}
colnames(sheet1_long) <- c("timestamp", "age", "going_out", "candy", "rating")
colnames(sheet1_long)

```

#Add column 'year'

```{r}
sheet1_long <- sheet1_long %>%
  add_column(year = 2015)
head(sheet1_long)
```


# Read in xlsx file

```{r}
sheet2 <- read_xlsx(here("raw_data/boing-boing-candy-2016.xlsx"))
sheet2
```

# Inspect the data i.e. dimensions, variable types, variable names, etc.?

```{r}
dim(sheet2)
colnames(sheet2)
glimpse(sheet2)
```

# Tidy data - long format

```{r}
sheet2_long <- sheet2 %>%
  pivot_longer(cols = starts_with("["),  #converting rows to columns 
               names_to = "candy",
               values_to = "rating")

dim(sheet2_long)
colnames(sheet2_long)
```

#Omitting columns that are not required

```{r}
sheet2_long <- sheet2_long %>%
  select(-c(6:22))

colnames(sheet2_long)
dim(sheet2_long)
```


```{r}
colnames(sheet2_long) <- c("timestamp", "going_out", "gender", "age", "country", "candy", "rating")
colnames(sheet2_long)
```

#Add column 'year'

```{r}
sheet2_long <- sheet2_long %>%
  add_column(year = 2016)
head(sheet2_long)
distinct(sheet2_long, country)
```



# Read in xlsx file


```{r}
sheet3 <- read_xlsx(here("raw_data/boing-boing-candy-2017.xlsx"))
sheet3
```

# Inspect the data i.e. dimensions, variable types, variable names, etc.?

```{r}
dim(sheet3)
colnames(sheet3)
glimpse(sheet3)
```

# Tidy data - long format

```{r}
sheet3_long <- sheet3 %>%
  pivot_longer(cols = starts_with("Q6"),  #converting rows to columns 
               names_to = "candy",
               values_to = "rating")
colnames(sheet3_long)
```

#Omitting columns that are not required

```{r}
sheet3_long <- sheet3_long %>%
  select(-c(6:17))
colnames(sheet3_long)
dim(sheet3_long)
```


#Change col names

```{r}
colnames(sheet3_long) <- c("internal_id", "going_out", "gender", "age", "country", "candy", "rating")
colnames(sheet3_long)

sheet3_long
#summarise((count = sum(is.na(rating))))
```


#Add column 'year'

```{r}
sheet3_long <- sheet3_long %>%
  add_column(year = 2017)
head(sheet3_long)
distinct(sheet3_long, country)
```

#X X X X X X X X X X X X X X X X X X X X X X X X X X X X X X 

# Combine all sheets 

```{r}

candy_combined <- bind_rows(sheet1_long, sheet2_long, sheet3_long)

colnames(candy_combined)
dim(candy_combined)
head(candy_combined)
```

#check

```{r}
distinct(candy_combined, going_out)
```


#Changing column data type

```{r}
candy_combined$age <- as.integer(candy_combined$age)
candy_combined$year <- as.integer(candy_combined$year)
candy_combined$internal_id <- as.integer(candy_combined$internal_id)
head(candy_combined)

```

# remove age greater than 122

```{r}
candy_combined <- candy_combined %>%
  filter(age <= 122)

length(candy_combined$age)
distinct(candy_combined, age)
```

# Change string values to lower case

```{r}
candy_combined <- candy_combined %>%
  mutate_if(is.character, str_to_lower)

head(candy_combined)
unique(candy_combined$country)
country_distinct <- distinct(candy_combined, country)

```


# Cleaning candy column

```{r}
candy_combined <- candy_combined %>%
  mutate(candy = str_extract(candy, "[^\\[\\]]+"),
         candy = str_extract(candy, "[^q6 |]+[^|]+"))

```

#cleaning country column

```{r}
canada_error <- c("can", "canada`")
uk_error <- c("endland", "england", "scotland", "u.k.", "united kindom", "united kingdom")
usa_error <- c("'merica", "america", "merica", "murica", "murrika", "n. america", "new jersey", "new york", "north carolina", "pittsburgh", 
               "the united states", "the united states of america", "the yoo ess of aaayyyyyy", "trumpistan", "u s", "u s a", "u.s.", "u.s.a.",  
               "unhinged states", "unied states", "unite states", "united  states of america", "united sates", "united staes", "united state", 
               "united statea", "united stated", "united states", "united states of america",
               "united statss", "united stetes", "united ststes", "unites states", "units states", "us", "us of a", "usa usa usa", "usa usa usa usa",
               "usa usa usa!!!!", "usa!", "usa! usa!", "usa! usa! usa!", "usa!!!!!!", "usa? hard to tell anymore..", "usaa", "usas", "usausausa",   
               "ussa")
candy_combined <- candy_combined %>%
  mutate(country = ifelse(country %in% canada_error, "canada", country),
         country = ifelse(country %in% uk_error, "uk", country),
         country = ifelse(country %in% usa_error, "usa", country))

country_distinct <- distinct(candy_combined, country)

```



```{r}
candy_combines <- candy_combined %>%
  filter(candy != "cash, or other forms of legal tender")
```



#XXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXX

1.4.2 Analysis questions
1. What is the total number of candy ratings given across the three years. (number of candy ratings, not number of raters. Don’t count missing values)

```{r}
candy_combined %>%
  filter(!is.na(rating)) %>%
  summarise(total_rating_count = n())

#total_candy_ratings
```



2. What was the average age of people who are going out trick or treating and the average age of people not going trick or treating?

```{r}
candy_combined %>%
  select(going_out, age) %>%
  filter(!is.na(going_out)) %>%
  group_by(going_out) %>%
  summarise(average_age = mean(age))



```


3. For each of joy, despair and meh, which candy bar revived the most of these ratings?

```{r}
candy_combined %>%
  select(candy, rating) %>%
  filter(!is.na(rating)) %>%
  group_by(rating, candy) %>%
  summarise(count = n()) %>%
  arrange(desc(count)) %>%
  slice(seq_len(1))



```

```{r}
distinct(candy_combined, candy)
```


4. How many people rated Starburst as despair?

```{r}
candy_combined %>%
  select(candy, rating) %>%
  filter(candy == "starburst", rating == "despair") %>%
  group_by(rating) %>%
  summarise(starburst_count = n()) #%>%
  #arrange(desc(count)) #%>%
  #slice(seq_len(3))
```


For the next three questions, count despair as -1, joy as +1 and meh as 0.

5. What was the most popular candy bar by this rating system for each gender in the dataset?

```{r}
candy_combined %>%
  select(gender, candy, rating) %>%
  filter(!is.na(rating)) %>%
  mutate(new_rating = case_when(rating == "despair" ~ -1,
                                rating == "joy" ~ 1,
                                rating == "meh" ~ 0,)) %>%
  group_by(gender, candy) %>%
  summarise(count_popular = sum(new_rating)) %>%
  arrange(desc(count_popular)) %>%
  slice(seq_len(1))

```


6. What was the most popular candy bar in each year?

```{r}
candy_combined %>%
  select(year, candy, rating) %>%
  filter(!is.na(rating)) %>%
  mutate(new_rating = case_when(rating == "despair" ~ -1,
                                rating == "joy" ~ 1,
                                rating == "meh" ~ 0,)) %>%
  group_by(year, candy) %>%
  summarise(popular_candy = sum(new_rating)) %>%
  arrange(desc(popular_candy)) %>%
  slice(seq_len(1))
```


7. What was the most popular candy bar by this rating for people in US, Canada, UK and all other countries?


```{r}

```







TEST
XXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXX



```{r}
sheet3_long %>%
  mutate(new_candy3= str_extract(candy, "[^Q6 |]+[^|]+"))
new_candy3
```


```{r}
sheet2_long %>%
  mutate(new_candy2 = str_extract(candy, "[^\\[\\]]+"))
new_candy2
```



