---
title: "R Notebook"
output: html_notebook
---

```{r}
install.packages("here")
```

```{r}
library(readxl)
library(readr)
library(dplyr)
library(here)
library(tidyverse)
library(stringr)
```


```{r}
here()
```

# Read in csv file

```{r}
# read in clean data from clean_data folder
candy_clean <- read_csv(here("clean_data/candy_clean.csv"), guess_max = 507000)
candy_clean
```



1.4.2 Analysis questions

1. What is the total number of candy ratings given across the three years. (number of candy ratings, not number of raters. Don’t count missing values)

```{r}
candy_clean %>%
  filter(!is.na(rating)) %>%  #exclude NA values
  summarise(total_rating_count = n())

#total_candy_ratings
```



2. What was the average age of people who are going out trick or treating and the average age of people not going trick or treating?

```{r}
candy_clean %>%
  select(going_out, age) %>%
  filter(!is.na(going_out)) %>%   #exclude NA values
  group_by(going_out) %>%    
  summarise(average_age = mean(age))

```


3. For each of joy, despair and meh, which candy bar revived the most of these ratings?

```{r}
candy_clean %>%
  select(candy, rating) %>%
  filter(!is.na(rating)) %>%     #exclude NA values
  group_by(rating, candy) %>%   
  summarise(count = n()) %>%
  arrange(desc(count)) %>%
  slice(seq_len(1))    #get the first record of each rating

```


4. How many people rated Starburst as despair?

```{r}
candy_clean %>%
  select(candy, rating) %>%
  filter(candy == "starburst", rating == "despair") %>%
  group_by(rating) %>%
  summarise(starburst_count = n()) 
  #arrange(desc(count)) #%>%
  #slice(seq_len(3))
```


For the next three questions, count despair as -1, joy as +1 and meh as 0.

5. What was the most popular candy bar by this rating system for each gender in the dataset?

```{r}
candy_clean %>%
  select(gender, candy, rating) %>%
  filter(!is.na(rating)) %>%
  mutate(new_rating = case_when(rating == "despair" ~ -1,    #create a new column with numeraical ratings
                                rating == "joy" ~ 1,
                                rating == "meh" ~ 0,)) %>%
  group_by(gender, candy) %>%
  summarise(count_popular = sum(new_rating)) %>%
  arrange(desc(count_popular)) %>%
  slice(seq_len(1))   #get the first record of each rating

```


6. What was the most popular candy bar in each year?

```{r}
candy_clean %>%
  select(year, candy, rating) %>%
  filter(!is.na(rating)) %>%
  mutate(new_rating = case_when(rating == "despair" ~ -1,  #create a new column with numeraical ratings
                                rating == "joy" ~ 1,
                                rating == "meh" ~ 0,)) %>%
  group_by(year, candy) %>%
  summarise(popular_candy = sum(new_rating)) %>%
  arrange(desc(popular_candy)) %>%
  slice(seq_len(1))  #get the first record of each rating
```


7. What was the most popular candy bar by this rating for people in US, Canada, UK and all other countries?


```{r}
candy_clean %>%
  select(year, candy, rating, country) %>%
  filter(!is.na(rating)) %>%
  filter(country %in% c("usa", "uk", "canada")) %>%
  mutate(new_rating = case_when(rating == "despair" ~ -1,  #create a new column with numeraical ratings
                                rating == "joy" ~ 1,
                                rating == "meh" ~ 0,)) %>%
  group_by(country, candy) %>%
  summarise(popular_candy = sum(new_rating)) %>%
  arrange(desc(popular_candy)) %>%
  slice(seq_len(1))  #get the first record of each rating
```






