---
title: "R Notebook"
output: html_notebook
---
```{r}
install.packages("here")
```


```{r}
library(readr)
library(dplyr)
library(here)
library(tidyverse)
```


```{r}
here()
```


```{r}
# read in clean data from clean_data folder
decathlon <- read_csv(here("clean_data/decathlon_clean.csv"))
decathlon
```


#1.1.2 Analysis questions

#1. Who had the longest long jump seen in the data?

```{r}
longest_jump <- decathlon %>%
  select(name, events, result) %>% 
  filter(events == "long_jump") %>% 
  arrange(desc(result)) %>%
  head(1)   # display only the first row
longest_jump
```

#2. What was the average 100m time in each competition?

```{r}
avg_100m <- decathlon %>%
  select(competition, events, result)%>%
  filter(events == "x100m") %>%
  group_by(competition) %>%
  summarise(avg_100 = mean(result))
avg_100m
```


#3. Who had the highest total points across both competitions?

```{r}
highest_total_points <- decathlon %>%
  distinct(name, points, competition) %>% # there are 10 rows for each name so using distinct() to ger one row for each name
  group_by(name) %>%
  summarise(total = sum(points)) %>%
  arrange(desc(total)) %>% 
  head(1)
highest_total_points
```


#4. What was the shot-put scores for the top three competitors in each competition?

```{r}
top_three <- decathlon %>%
  select(name, events, result, competition) %>%
  filter(events == "shot_put") %>%
  group_by(competition) %>%
  arrange(desc(result)) %>%
  slice(seq_len(3)) #get desired length of the sequence that is grouped by competition
top_three
```

#5. What was the average points for competitors who ran the 400m in less than 50 seconds vs. those than ran 400m in more than 50 seconds?


```{r}
average_points <- decathlon %>%
  select(name, events, result) %>%
  filter(events == "x400m") %>%
  mutate(threshold_above_50 = result > 50) %>% #create a new column which stores boolean result based on condition specified
  group_by(threshold_above_50) %>% #using the new column to group by
  summarise(average_400m = mean(result))

average_points
```









